Quality Criteria for SKOS Vocabularies

### 2012-03-22 criteria.tex
added comments/updated result table

This criterion is an extension of a statement from \cite{Isaac2009} that recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. Based on our review of existing vocabularies, we could observe that (i) many vocabularies don’t use concept schemes and (ii) identical \texttt{skos:altLabels} or \texttt{skos:hiddenLabels} can also negatively affect use cases such as auto-completion, which proposes labels based on the user input. Thus, this criterion generalizes this statement by finding all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:ULV \times ULV \rightarrow [0,1]$. The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. This criterion is calculated by iterating over all pairs of concepts, i.e., elements of $C \times C$, applying $sim$ to every possible label combination, and collecting those pairs for which $sim$ meets or exceeds a specified threshold.

@@

\subsubsection{Top Concept Having Broader Concepts}
\cite{Allemang2011} proposes to ``not indicate any concepts internal to the tree as top concepts'', i. e. top concepts should not have broader concepts. This criteria identifies all elements from $C$ that occur in both $I_{EXT}(\texttt{skos:topConceptOf}^\mathcal{I})$ and as first element of each pair from $I_{EXT}(\texttt{skos:broader}^\mathcal{I})$.

@@

\subsubsection{Missing In-Links}
A periodically updated version of a diagram visualizing ``the data sets in the LOD cloud as well as their interlinkage relationships'' is publised online\footnote{\url{http://www4.wiwiss.fu-berlin.de/lodcloud/state/}}. In these diagrams, dbpedia is shown as the most intensely linked resource, i.e. a large number of vocabularies link to resources in that dataset. Therefore a dataset with a large number of other datasets referencing it, can be assumed to be of great value for the community. Accordingly, estimating the number of In-Links of SKOS concepts defined in a vocabulary (henceforth called the in-degree of a concept) would give some hint about how established or accepted a controlled vocabulary is. A method to carry out these estimations would be utilizing existing Linked Data indices like, e.g., Sindice\footnote{\url{http://sindice.com/}} that provide a SPARQL endpoint. In order to estimate the number of In-Links for a conceptual resource, we iterate over all elements in $C$ and query the Sindice SPARQL endpoint for triples containing the concept's URI in the object part. Those concepts having an empty query result, are missing any In-Links and are thus included in the result set of this criterion.

\subsubsection{Missing Out-Links}
The Web of Data consists of globally interconnected datasets, ``enabling seamless connections between data sets''\cite{Heath2011}. These links can be established between various SKOS vocabularies by Out-Links, i.e., references from conceptual resources in the ``source vocabulary'' $V$ to resources identified by an HTTP URI and available at a different host. Similar to the criterion above, this criterion identifies the set of all conceptual resources that lack such Out-Links. Calculation of the result set is done by iterating over all elements $c \in C$ and investigating the host part of all related resources. If none of these host names differs from the host part of $c$, it has no Out-Links and is thus included into the resources identified by this criterion.

@@

\subsubsection{Link Target Unavailability}
This criterion identifies a set of all resources that have been identified as unavailable, i.e. resolving their URI leads to an erroneous HTTP response or no response at all. An erroneous HTTP response in that case can be defined as a response code (after a possible redirection) other than 200. Just as in the ``traditional'' Web, these ``broken links'' hinder information gathering and should be avoided. Calculation of this criterion is performed by iterating over all resources in $IR$, performing an HTTP request and include those resources in the result set that are unavailable.

>>>>

This criterion is an extension of a statement from \cite{Isaac2009} that recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. Based on our review of existing vocabularies, we could observe that (i) many vocabularies don’t use concept schemes and (ii) identical alternative or hidden can also negatively affect use cases such as auto-completion, which proposes labels based on the user input. Thus, this criterion generalizes this statement by finding all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:ULV \times ULV \rightarrow [0,1]$. The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. This criterion is calculated by iterating over all pairs of concepts, i.e., elements of $C \times C$, applying $sim$ to every possible label combination, and collecting those pairs for which $sim$ meets or exceeds a specified threshold.

@@

\paragraph{\textbf{QF 11: Top Concept Having Broader Concepts}}
\cite{Allemang2011} proposes to ``not indicate any concepts internal to the tree as top concepts'', i. e. top concepts should not have broader concepts. This criteria iterates over all \texttt{skos:TopConcept}s defined in $V$ and collects those that are related to a resource by the \texttt{skos:broader} property.

@@

\subsubsection{Missing In-Links}
A periodically updated version of a diagram visualizing ``the data sets in the LOD cloud as well as their interlinkage relationships'' is publised online\footnote{\url{http://www4.wiwiss.fu-berlin.de/lodcloud/state/}}. In these diagrams, DBpedia is shown as the most intensely linked resource, i.e. a large number of vocabularies link to resources in that dataset. This indicates that a dataset with a large number of other datasets referencing it, can be assumed to be of great value for the community, adding cohesion to the LOD cloud. Accordingly, estimating the number of In-Links of SKOS concepts defined in a vocabulary (henceforth called the in-degree of a concept) gives some hint about how established or accepted a controlled vocabulary is. A method to carry out these estimations would be utilizing existing Linked Data indices like, e.g., Sindice\footnote{\url{http://sindice.com/}} that provide a SPARQL endpoint. In order to estimate the number of In-Links for a conceptual resource, we iterate over all elements in $C$ and query the Sindice SPARQL endpoint for triples containing the concept's URI in the object part. Those concepts having an empty query result, are missing any In-Links and are thus included in the result set of this criterion.

\subsubsection{Missing Out-Links}
The Web of Data consists of globally interconnected datasets, ``enabling seamless connections between data sets''\cite{Heath2011}. These links can be established between various SKOS vocabularies by Out-Links, i.e., references from conceptual resources in the ``source vocabulary'' $V$ to resources identified by an HTTP URI and available at a different host. Similar to the criterion above, this criterion identifies the set of all conceptual resources that lack such Out-Links. Calculation of the result set is done by iterating over all elements $c \in C$ and investigating the host part of all related resources. If none of these host names differ from the host part of $c$, it has no Out-Links and is thus included into the resources identified by this criterion.

@@

\subsubsection{Link Target Unavailability} \todo{BH}{should we just call this "Broken Links" because that's essentially what it's all about}
This criterion identifies a set of all resources that have been identified as unavailable, i.e. resolving their URI leads to an erroneous HTTP response or no response at all. An erroneous HTTP response in that case can be defined as a response code (after a possible redirection) other than 200. Just as in the ``traditional'' Web, these ``broken links'' hinder information gathering and should be avoided. Calculation of this criterion is performed by iterating over all resources in $IR$, performing an HTTP request and include those resources in the result set that are unavailable.
